---
title: "car_sales_project"
author: "SILAS"
date: "2024-10-10"
output: html_document
---
# Read the CSV file into R
# View the first few rows of your data
```{r}
car_sales <- read.csv("C:/Users/Administrator/Desktop/car_sales.csv")
head(data)
```
## In order to start cleaning the data, i will need to install the required packages. like `tidyverse`, `skimr`, and `janitor` and `dplyr`.
```{r}
install.packages("tidyverse")
library(tidyverse)

install.packages("skimr")
library(skimr)

install.packages("here")
library("here")

install.packages("janitor")
library(janitor)
library("janitor")

install.packages("dplyr")
library(dplyr)
```
# Summary statistics,summary(data)
# View the data structure,str(data)
# Check for missing values,colSums(is.na(data))
# View the first few rows, head(data)
## Step 3: Exploring or getting to know the data, 
```{r}
head(car_sales)
str(car_sales)
glimpse(car_sales)
summarise(car_sales)
colnames(car_sales)
colSums(is.na(car_sales))

skim_without_charts(car_sales)
```
```{r}
# Converting columns to appropriate data types if necessary
car_sales$Sales_in_thousands <- as.numeric(car_sales$Sales_in_thousands)
car_sales$Price_in_thousands <- as.numeric(car_sales$Price_in_thousands)
car_sales$Fuel_efficiency <- as.numeric(car_sales$Fuel_efficiency)

```
```{r}
# exploring the data mor
# Calculating the mean, median, and standard deviation for sales and price
mean(car_sales$Sales_in_thousands, na.rm = TRUE)
median(car_sales$Price_in_thousands, na.rm = TRUE)
sd(car_sales$Fuel_efficiency, na.rm = TRUE)

```
```{r}
# Correlation matrix of relevant numeric variables
numeric_vars <- car_sales[, c("Sales_in_thousands", "Price_in_thousands", "Horsepower", "Fuel_efficiency", "Curb_weight")]
cor(numeric_vars, use="complete.obs")

```
```{r}
# Scatterplot of Price vs Sales
plot(car_sales$Price_in_thousands, car_sales$Sales_in_thousands,
     xlab = "Price in Thousands", ylab = "Sales in Thousands",
     main = "Sales vs Price", col = "blue", pch = 19)

# Boxplot: Vehicle type vs Price
boxplot(Price_in_thousands ~ Vehicle_type, data = car_sales, 
        main="Price by Vehicle Type", xlab="Vehicle Type", ylab="Price in Thousands")

```
```{r}
# Histogram of Horsepower
hist(car_sales$Horsepower, main="Distribution of Horsepower", xlab="Horsepower", col="green")

# Histogram of Fuel Efficiency
hist(car_sales$Fuel_efficiency, main="Distribution of Fuel Efficiency", xlab="Fuel Efficiency (mpg)", col="orange")

```
```{r}
# Linear regression model to predict Sales
lm_model <- lm(Sales_in_thousands ~ Price_in_thousands + Horsepower + Fuel_efficiency, data = car_sales)
summary(lm_model)

# Plot the residuals of the model
plot(lm_model)

```
```{r}
# Installing and loading the rpart package for decision trees
install.packages("rpart")
library(rpart)

# Decision tree to predict vehicle type based on numeric features
tree_model <- rpart(Vehicle_type ~ Price_in_thousands + Horsepower + Fuel_efficiency, data = car_sales)
plot(tree_model)
text(tree_model, use.n = TRUE)

```

