---
title: "Titanic Survival Analysis"
author: "SILAS"
date: "2024-10-13"
output: html_document
---
First, let us install the pacakges
```{r}
install.packages("dplyr")
install.packages("ggplot2")
install.packages("tidyverse")
```


let us load those packages as well
```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
```


step two , loading the dataset (train.csv)
```{r}
train<- read_csv("train.csv")
```


 step 3: inspecting the dataset
```{r}
head(train)
view(train)
str(train)
colnames(train)
names(train)
summary(train)
```

Let us check for missing values
```{r}
missing_value<- is.na(train)
missing_value
```

We can use this also,which will return the number of NA in each column
```{r}
colSums(is.na(train))
```

let remove the NA in age
```{r}
mean(train$Age,na.rm = TRUE)
```

To find the number of NA in each colunms of train 
```{r}
colSums(is.na(train))
```

to imput the missing numbers using median
```{r}
train$Age[is.na(train$Age)] <- median(train$Age, na.rm = TRUE)
```


 Fill missing Embarked values with the most frequent value "s"
```{r}
train$Embarked[is.na(train$Embarked)] <- "S"
```
 
To find the maximum,min, mean,mode and meadian age in the train
```{r}
max(train$Age,na.rm = TRUE)
min(train$Age,na.rm = TRUE)
mean(train$Age,na.rm = TRUE)
median(train$Age,na.rm = TRUE)
```

To select Age, Name,Sex
```{r}
data<-train %>%
  select(Age,Name,Sex) %>%
  filter(Age>30 & Sex == "male")
view (data)
```

TO FIND THE SURVIVAL RATE BY AGE, PCLASS,SEX
```{r}
# Survival rate by Gender
survival_age<- train %>%
               group_by(Sex) %>%
               summarize(Survival_Rate = mean(Survived))

view(survival_age)
# Survival rate by Pclass
train %>%
  group_by(Pclass) %>%
  summarize(Survival_Rate = mean(Survived))

# Survival rate by Age Group
train %>%
  mutate(AgeGroup = cut(Age, breaks = seq(0, 80, by = 10))) %>%
  group_by(AgeGroup) %>%
  summarize(Survival_Rate = mean(Survived))

# Survival rate by Embarked
train %>%
  group_by(Embarked) %>%
  summarize(Survival_Rate = mean(Survived))

```

TO VISUALIZE THE DATA
```{r}
ggplot(train, aes(x = factor(Survived))) +
  geom_bar(fill = "steelblue") +
  labs(x = "Survived", y = "Count", title = "Survival Count")
```

OR I PREFER THIS SYNTAX BELOW

```{r}
# Load ggplot2 library
library(ggplot2)

# Create the bar plot
ggplot(data = train) +
  geom_bar(mapping = aes(x = factor(Survived)), fill = "steelblue") +
  labs(
    title = "Survival Count",      # Title of the plot
    subtitle = "man",              # Subtitle of the plot
    caption = "titanic",           # Caption of the plot
    x = "Survived"                 # Label for the x-axis
  )

```
ox plot for Age and Survival

```{r}
library(ggplot2)

ggplot(data = train) +
  geom_histogram(mapping = aes(x = Age), fill = "steelblue", bin = 30, color = "black") +
  labs(
    title = "Age Distribution by Survival",
    subtitle = "Titanic Dataset",
    caption = "Source: Titanic",
    x = "Age",
    y = "count"
  )

```


PLOTING HISTOGRAM FOR AGE DISTRIBUTION
```{r}
ggplot(data = train) +
  geom_boxplot(mapping = aes(x = factor(Survived), y = Age), fill = "steelblue") +
  labs(
    title = "Age Distribution by Survival",
    subtitle = "Titanic Dataset",
    caption = "Source: Titanic",
    x = "Survived",
    y = "Age"
  )

```

```{r}
# Bar plot for Gender and Survival

library(ggplot2)

ggplot(data = train) +
  geom_bar(mapping = aes(x = Sex, fill = factor(Survived))) +
  labs(
    title = "Survival by Gender",
    subtitle = "Titanic Dataset",
    caption = "Source: Titanic",
    x = "Sex",
    y = "Count"
  )

```